### 2022-04-18 paper/localpsf.tex
deleted old stuff from paper

\subsubsection{Improvement for symmetric operators}
\label{sec:symmetric_improvements}
If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$. Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$. We could form an alternative approximation of $\Phi(y,x)$ by performing the same interpolation as before, but with $x$ and $y$ taking opposite roles. Both of these interpolation schemes can be combined to improve the approximation accuracy at negligible additional cost. We combine these interpolation schemes by placing function values 
\begin{equation*}
\left(\widehat{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $x_i-x$, placing function values
\begin{equation*}
\left(\widehat{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.

In detail, define 
\begin{equation*}
\widehat{x_i} := x_i - x \qquad \text{and} \qquad \widehat{y_i} := y_i - y.
\end{equation*}
and let
\begin{equation*}
\widehat{B} = \begin{bmatrix}
\widehat{B}_{xx} & \widehat{B}_{xy} \\
\widehat{B}_{yx} & \widehat{B}_{xx}
\end{bmatrix}, \qquad \widehat{w} = \begin{bmatrix}\widehat{w}_x \\ \widehat{w}_y\end{bmatrix}, \qquad \widehat{f} = \begin{bmatrix}
\widehat{f}_x \\ \widehat{f}_y
\end{bmatrix},
\end{equation*}
be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ given by
\begin{align*}
\left(\widehat{B}_{xx}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{x}_j\|\right), &\qquad
\left(\widehat{B}_{xy}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{y}_j\|\right), \\
\left(\widehat{B}_{yx}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{x}_j\|\right), &\qquad \left(\widehat{B}_{yy}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{y}_j\|\right),
\end{align*}
for $i,j=1,\dots,k$. We approximate $\Phi(y,x)$ as follows:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widehat{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widehat{y}_i\|\right),
\end{equation}
where $\widehat{w}$ is the solution of the following regularized least-squares problem,
\begin{equation*}
\min_{\widehat{w}} \frac{1}{2}\|\widehat{B} \widehat{w} - \widehat{f}\|^2 + \frac{\alpha}{2}\|\widehat{w}\|^2,
\end{equation*}
for a small regularization parameter $\alpha$. 
We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$. We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice. In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.

@@

\subsection{Rational positive semi-definite modification}
\label{sec:make_spd}

@@

\section{old rational}

In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.

Because of the modified interpolation for symmetric matrices described in Section \ref{sec:symmetric_improvements}, $\mathbf{A}_H$ is nearly (but not completely) symmetric. We further symmetrize $\mathbf{A}_H$ as follows:
\begin{equation*}
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
\end{equation*}
This step is fast because adding H-matrices is a comparatively inexpensive operation.

We modify $\mathbf{A}_H^{\text{sym}}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
\begin{equation}
\label{eq:rational_matrix_function}
\mathbf{A}_H^{\text{sym}+} := \ratcoeff_0 \bm{I} + \ratcoeff_1 \mathbf{A}_H^{\text{sym}} + \ratcoeff_2 \left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1},
\end{equation}
where $\bm{I}$ is the identity matrix which has the same shape as $\mathbf{A}_H^{\text{sym}}$. Approximation \eqref{eq:rational_matrix_function} may be written as $\mathbf{A}_H^{\text{sym}+} = \ratfct(\mathbf{A}_H^{\text{sym}})$, where $\ratfct$ is the following scalar rational function:
\begin{equation}
\label{eq:rational_function_scalar}
\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole},
\end{equation}
which is extended to matrices by functional calculus.
The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}+}$ approximately equal the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$, while the negative and small positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are modified so that $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite. Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\mathbf{A}_H^{\text{sym}+}$. In Figure FIG we illustrate $\ratfct(\lambda)$. Forming $\mathbf{A}_H^{\text{sym}+}$ requires computing an H-matrix inverse, $\left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1}$. While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}$ is an H-matrix. Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.

We choose $\ratcoeff_0, \ratcoeff_1, \ratcoeff_2$, and $\ratpole$, to be the solution to the following system of equations:
\begin{subequations}
	\label{eq:four_eq_system}
	\begin{align}
	\ratfct(0) &= 0 \label{eq:desired_property_1} \\
	\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
	\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
	\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
	\end{align}
\end{subequations}
where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\mathbf{A}_H^{\text{sym}}$, respectively. We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}. We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\mathbf{A}_H^{\text{sym}}$ is already positive semi-definite, so we do not need to modify it with a rational function. We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$. We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}. Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.

Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite. Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\mathbf{A}_H^{\text{sym+}}$ contains the null-space of $\mathbf{A}_H^{\text{sym}}$. This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero. Hessians in ill-posed distributed parameter inverse problems typically have both of these properties. Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ to be close to the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$. I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\mathbf{A}_H^{\text{sym}}$ as little as possible. Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$. There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ will be closer to the corresponding positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ (good), but the erroneous negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ will be amplified in magnitude (bad). This is shown in Figure FIG. We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.

\subsection{Solving for rational function scalars}
\label{app:solve_rational_system}

Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
\begin{equation}
	\label{eq:rational_3_linear_system}
	\begin{bmatrix} 
		1 & 0 & 1/\ratpole \\ 0 & 1 & -1/\ratpole^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\ratpole) 
	\end{bmatrix}
	\begin{bmatrix} 
		\ratcoeff_0 \\ \ratcoeff_1 \\ \ratcoeff_2 
	\end{bmatrix}
	=
	\begin{bmatrix} 
		0 \\ 0 \\ \lambda_\text{max} 
	\end{bmatrix},
\end{equation}
and we may write Equation \eqref{eq:desired_property_3} as follows: 
\begin{equation}
	\label{eq:rational_extra_eq}
	\ratcoeff_0 + \ratcoeff_1 \lambda_\text{min} + \frac{\ratcoeff_2}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}| = 0.
\end{equation}
Let 
\begin{equation*}
	f(\ratpole) := \ratcoeff_0(\ratpole) + \ratcoeff_1(\ratpole) \lambda_\text{min} + \frac{\ratcoeff_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
\end{equation*}
denote the left hand side of \eqref{eq:rational_extra_eq}, but with $\ratcoeff_0(\ratpole)$, $\ratcoeff_1(\ratpole)$, and $\ratcoeff_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$. I.e., within $f(\ratpole)$ the quantities $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system. We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
\begin{equation}
	\label{eq:g_mu_zero}
	f(\ratpole)=0
\end{equation}
with bound $\ratpole > |\lambda_\text{min}|$ and initial guess $\ratpole_0 = 2 |\lambda_\text{min}|$. Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\ratpole$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$.

\section{Proofs}
\label{app:proofs}

\begin{defn}[Moments of $\impulseresponse_x$]
	\label{defn:alpha_rho_mu_sigma}
	We define the spatially varying scaling factor, $\spatialvol$, the scaled impulse response, $\widehat{\impulseresponse}_x$, the spatially varying mean, $\spatialmean$, and the spatially varying covariance, $\spatialcov$, as follows:
	\begin{align*}
		\spatialvol:\Omega \rightarrow \mathbb{R}, &\qquad \spatialvol(x) := \int_{\Omega} \impulseresponse_x(z) dz, \\
		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \spatialvol(x), \\
		\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim, &\qquad \spatialmean(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
		\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}, &\qquad\spatialcov(x) := \int_\Omega (z - \spatialmean(x))(z - \spatialmean(x))^T \widehat{\impulseresponse}_x(z) dz.
	\end{align*}
\end{defn}

\begin{proof}[Proof of Proposition \ref{thm:vol_mean_cov}]
	Let $v \in L^2$ be arbitrary. We have
	\begin{align*}
		\left(v, \spatialvol\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \left(\Aop \delta_x^*\right)^*(z) dz dx\\
		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
		=& \int_\Omega \int_\Omega v(x) A(z,x) C(z) dz dx \\
		=& \left(\Aop^T C\right)(v),
	\end{align*}
	which implies $\spatialvol = \left(\Aop^T C\right)^*$.

	Using similar techniques, we have
	\begin{align*}
		\left(v, \spatialvol \cdot \spatialmean^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_{\Omega} z^i \left(\Aop \delta_x^*\right)^*(z)/\spatialvol(x) dz dx\\
		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
		=& \left(\Aop^T L^i\right)(v),
	\end{align*}
	which implies $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$.

	Using the identity
	\begin{equation*}
		\spatialcov(x) = \int_\Omega zz^T \widehat{\impulseresponse}_x(z) dz - \spatialmean(x) \spatialmean(x)^T,
	\end{equation*}
	we have
	\begin{align*}
		\left(v, \spatialvol \cdot \left(\spatialcov^{ij}+\spatialmean^i \cdot \spatialmean^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_\Omega z^i z^j \left(\Aop \delta_x^*\right)^*(z) / \spatialvol(x) dz dx \\
		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
		=& \left(\Aop Q^{ij}\right)(v),
	\end{align*}
	which implies $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot\spatialmean^j$.
\end{proof}

\section{Background: distributions}
\label{app:distributions}

%Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. 
The action of $\Aop$ is extended to distributions $\rho$
%$\genericdistribution:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ 
via the formula
\begin{equation*}
	\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy, 
\end{equation*}
where $\Phi(y,\cdot)$ is the function $x \mapsto \Phi(y,x)$. This is derived formally as follows:
\begin{align*}
	\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx dy \\
	&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx \\
	&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
\end{align*}
For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
\begin{equation}
	\label{eq:appendix_impulse_response}
	\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
\end{equation}
Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$. That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.

\section{Discretization}
\label{sec:discretization}

Let $\febasis_i$, $i=1,\dots,\fedim$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
\begin{equation*}
	V_h = \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)
\end{equation*}
be the corresponding  finite element space with the $L^2$ inner product. Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$. In turn, functions $u_h\in V_h$ are represented in computations by length $\fedim$ arrays $\bm{u}$, such that the array entries of $\bm{u}$ are the coordinates of $u_h$ in the finite element basis. That is,
\begin{equation}
	\label{eq:finite_bijection}
	u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}

We write $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ to denote the finite element \emph{mass matrix}, which has entries
\begin{equation*}
	\bm{M}_{ij} = \left(\febasis_i, \febasis_j\right)_{L^2(\Omega)} = \int_{\Omega} \febasis_i(x) \febasis_j(x) dx,
\end{equation*}
and we write $\mathbb{R}^\fedim_{\bm{M}}$ to denote the space $\mathbb{R}^\fedim$ with the matrix-weighted inner product
\begin{equation*}
	\left(\bm{u}, \bm{v}\right)_{\bm{M}} := \bm{v}^T \bm{M} \bm{u}.
\end{equation*}
Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^\fedim_{\bm{M}}$. That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\bm{u},\bm{v}\right)_{\bm{M}}$. Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.

Here vectors in $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$ and $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ are viewed as column vectors, and $\bm{u}^T$ and $\bm{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\bm{u}$ and $\bm{\sigma}$, respectively. The discrete dual of a vector $\bm{v}$ with respect to the $\bm{M}$ inner product is written as $\bm{v}^*$, and $\bm{v}^*$ is viewed as a column vector.

\subsection{Discretized operations}
\label{app:discretized_operations}

Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^\fedim_{\bm{M}}$ and $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.

\begin{description}
	\item[Coefficients of a functional:] The coefficient dual vector, $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
	\begin{equation*}
		\bm{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,\fedim.
	\end{equation*}
	\item[Action of a functional:] The action of a linear functional $\sigma_h \in V_h'$ on a function $u_h \in V_h$ may be computed as follows:
	\begin{equation*}
		\sigma_h(v_h) = \bm{\sigma}^T \bm{u}.
	\end{equation*}
	\item[Riesz representation:] Let $\sigma_h \in V_h'$. The Riesz representation of $\sigma_h$ is the unique  function $\sigma_h^* \in V_h$ satisfying 
	\begin{equation*} 
		\left(\sigma_h^*, v_h\right)_{L^2(\Omega)} = \sigma_h(v_h) \quad\forall v_h \in V_h.
	\end{equation*}
	The coefficient dual vector $\bm{\sigma}^*$ corresponding to $\sigma_h$ is given by
	\begin{equation*} 
		\bm{\sigma}^* = \bm{M}^{-1} \bm{\sigma}.
	\end{equation*}
	\item[Finite element $L^2$ projection:] Let $g \in L^2(\Omega)$. The orthogonal projection of $g$ onto $V_h$ with respect to the $L^2$ inner product is the unique function $g^\pi_h \in V_h$ satisfying
	\begin{equation*}
		\left(g^\pi_h, v_h\right)_{L^2(\Omega)} = \left(g, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h.
	\end{equation*}
	 Let $\left(\bm{g}^\pi\right)^* \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ be the vector with components
	\begin{equation*}
		\left(\bm{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,\fedim.
	\end{equation*}
	The coefficient vector $\bm{g}^\pi \in \mathbb{R}^\fedim_{\bm{M}}$ corresponding to $g^\pi_h$ is given by
	\begin{equation*}
		\bm{g}^\pi = \bm{M}^{-1}\left(\bm{g}^\pi\right)^*.
	\end{equation*}
	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots,\fedim$, be the Lagrange nodes associated with the finite element basis functions $\febasis_i$. The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
	\begin{equation*}
		g_h^I(x) = \sum_{i=1}^\fedim g(p_i) \febasis_i(x).
	\end{equation*}
	The coefficient vector, $\bm{g}$, for $g_h^I$ has components 
	\begin{equation*}
	\bm{g}_i^I = g(p_i), \quad i=1.,\dots,\fedim.
	\end{equation*}
	\item[Projection vs. interpolation:]
	Typically, $g^\pi$ is a more accurate approximation of $g$ than $g^I$, but $g^I$ is cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not. If the finite element basis satisfies the following Kronecker property,
	\begin{equation*}
		\febasis_i(p_j) = \begin{cases}1 & i=j \\
		0 & \text{otherwise},\end{cases}
	\end{equation*}
	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,\fedim$ (i.e., the interpolant equals the original function at the Lagrange nodes). Interpolation is typically sufficiently accurate for practical purposes if the Kronecker property is satisfied. Projection should be used if the Kronecker property is not satisfied.
	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\bm{B} \in \bm{R}^{\fedim \times \fedim}$ be the matrix with entries
	\begin{equation*}
		\bm{B}_{ij} = \left(\mathcal{B}_h \febasis_j\right)(\febasis_i).
	\end{equation*}
	The matrix $\bm{B}$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ by matrix multiplication, and
	\begin{equation*}
		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{u}.
	\end{equation*}
	\item[Transpose:] The matrix $\bm{B}^T$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ via matrix multiplication, and
	\begin{equation*}
		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B}^T \bm{u}.
	\end{equation*}

	\item[Distributions:] Let $\genericdistribution:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution. If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,\fedim$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$. The functional $\genericdistribution_h$ has a coefficient dual vector $\bm{\genericdistribution} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ with entries 
	\begin{equation*}
		\bm{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,\fedim.
	\end{equation*} 
	For example, let $\delta_p$ be the Dirac distribution centered at a point $p\in\Omega$. If the finite element basis functions, $\febasis_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$. The corresponding coefficient dual vector, $\bm{\delta}_p$, has entries 
	\begin{equation*}
		\left(\bm{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,\fedim.
	\end{equation*}

	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\genericdistribution_h \in V_h$, of a distribution, $\genericdistribution$, to the space $V_h$ has a a Riesz representation $\genericdistribution_h^* \in V_h$. Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$. For $v_h \in V_h$, we have
	\begin{align*}
		\langle \mathcal{B}_h, \genericdistribution_h \rangle(v) &= \int_\Omega v_h(y) \genericdistribution_h\left(B_h(y, \cdot)\right) dy \\
		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \genericdistribution_h^*(x) dx\right) dy \\
		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \genericdistribution_h^*(x) dx dy = \left(\mathcal{B}_h \genericdistribution_h^*\right)(v).
	\end{align*}
	Going from the first to the second line we used the definition of the Riesz representation. The action of $\mathcal{B}_h$ on $\genericdistribution_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\genericdistribution^*_h$ in the conventional sense. We have
	\begin{equation*}
		\sigma_h = \langle \mathcal{B}_h, \genericdistribution_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \genericdistribution_h^* \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{M}^{-1} \bm{\genericdistribution}.
	\end{equation*}
\end{description}

\subsection{Discretized mean and covariance estimation}

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Operator $\bm{A}$}
		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}

		$\bm{\spatialvol} = \left(\bm{M}^{-1}\bm{A}^T \bm{C}\right)^*$

			$\bm{\spatialmean}^i = \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right)^* / \bm{\spatialvol}$
		}
		\For{$i=1,2,\dots,\gdim$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $\bm{Q}^{ij} = x^i x^j$

				$\bm{\spatialcov}^{ij} = \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right)^* / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$

				$\bm{\spatialcov}^{ij} = \bm{\spatialcov}^{ji}$

	}
	\caption{Discretized version of Algorithm \ref{alg:varhpi_mean_cov}}
	\label{alg:varhpi_mean_cov_discrete}
\end{algorithm2e}

In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\bm{\spatialvol}, \bm{\spatialmean}^{ij}, \bm{\spatialcov}^{ij}\in\mathbb{R}^\fedim_{\bm{M}}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property). In our numerical examples we use finite element basis functions that satisfy the Kronecker property, and we use finite element interpolations. The coefficient vectors corresponding to either the interpolations or projections are denoted $\bm{C}$, $\bm{L}^i$, and $\bm{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
	\bm{\spatialvol} &= \bm{M}^{-1}\bm{A}^T\bm{C} \\
	\bm{\spatialmean}^i &= \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}\\
	\bm{\spatialcov}^{ij} &= \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j,
\end{align*}
respectively. Here $\bm{f} \cdot \bm{g}$ and $\bm{f} / \bm{g}$ denote the component-wise multiplication and division of vectors, respectively.

\subsection{Discretized sampling of impulse responses}
In computations, the Dirac comb coefficient dual vector for $\diraccomb_h^k$ is the vector $\bm{\diraccomb}^k \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ which has entries
\begin{equation*}
	\bm{\diraccomb}^k_i = \sum_{x_j \in \pointbatch_k} \febasis_i(x_j), \quad i=1,\dots,\fedim.
\end{equation*}
The Dirac comb impulse response coefficient vector for $\combresponse^h_k$ is the vector $\bm{\combresponse}_k \in \mathbb{R}^\fedim_{\bm{M}}$ given by
\begin{equation*}
	\bm{\combresponse}_k = \bm{M}^{-1} \bm{A}^T \bm{M}^{-1}\bm{\diraccomb}^k.
\end{equation*}

\section{OLD CONTENT}

\subsection{disorganized march 21 22}

H-matrix approximations are highly desirable because

The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
Classical H-matrix construction methods 
require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points. We overcome this problem by c

We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source.

We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed).

access to $\Aop$ is available through ``operator actions.'' That is,

After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored. This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$. However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.

For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.

for a collection of points $x_i$ scattered throughout $\Omega$.
Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
\begin{equation}
%\label{eq:impulse_response_defn}
\impulseresponse_x(y) := \Aker(y, x),
\end{equation} 
be the \emph{impulse response} of $\Aop$ at $x$.

Because we build a H-matrix, and

>>>>

@@

\section{Rational positive semi-definite modification}
\label{sec:make_spd}

@@
